Modal Verification Patterns for Systems Software

### 2025-09-15 paper.tex
a couple live fixes during review

% \subsection{Conclusion \& Contributions}
% \paragraph{Conclusion} 
Interacting with computing resources is the essential point which shapes the design of low-level systems software. 
Exploiting these design choices has been an important field of study.
This survey examines these common patterns in systems software, and their relationships
to modalities. We believe that certain properties of modalities enable us to understand and 
address the verification challenges of these systems by tailoring specification and proof
to the same patterns used to design the systems.

@@

Newer work surveyed here adapts this idea to deal with different structures specified in

@@

leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory. DAG abstraction instead of recursive decent tree allows non-duplication on the reference counts for the identical memory objects.

@@

On the other hand, program logics such as \textsf{Iris}~\cite{jung2018iris} utilizes Kripke model in ordering the separation algebras (a collection of \textsc{CMRA}s as an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}) both by heap extension and step index. 
This is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{nakano-00}.
However in general, many modalities can be given proof-theoretic semantics directly within a logic. 
This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} 
logic itself) , as well as for a wide range of other modalities in the 
literature~\cite{restall2002introduction}. These modalities are typically multi-modal, 
with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$.

@@

This definition also reveals a subtlety about the denotation of assertions $\phi$ --- they are effectively \emph{predicates} of some kind, functions from some possible world into some logical algebra such as a BI-algebra~\cite{ohearn1999bunched}. 
Taking this interpretation literally gives rise to a convenient way to embed new modalities in \textsf{Iris} 
in a way well-suited to many use cases. 
If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the \emph{pointwise lifting} of the algebra. A number of 
projects for for systems software verification now take advantage of this,
explicitly or implicitly lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates 
used as assertions within modalities. 
Connecting this to \textsf{Iris}'s typeclasses for interactive proofs~\cite{krebbers2018mosel,krebbers-17} then permits working within the embedded logic as in \textsf{Iris}.

>>>>

% \subsection{Conclusion \& Contributions}
% \paragraph{Conclusion} 
Interacting with computing resources is the essential point which shapes the design of low-level systems software. 
Exploiting similar design choices has been an important field of study.
This survey examines these common patterns in systems software, and their relationships
to modalities. We believe that certain properties of modalities enable us to understand and 
address the verification challenges of these systems by tailoring specification and proof
to the same patterns used to design the systems.

@@

Newer work~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024}:w
 surveyed here adapts this idea to deal with different structures specified in

@@

leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory. DAG abstraction instead of recursive descent tree allows non-duplication on the reference counts for the identical memory objects.

@@

On the other hand, program logics such as \textsf{Iris}~\cite{jung2018iris} utilizes Kripke model in ordering the separation algebras (a collection of \textsc{CMRA}s as an algebraic cousin of a BI-algebra~\cite{ohearn1999bunched}) both by heap extension and step index. 
This is the case for \textsf{Iris}'s one built-in modality for step-indexing~\cite{nakano-00}.
However in general, many modalities can be given proof-theoretic semantics directly within a logic. 
This is the case for \textsf{Iris}'s weakest precondition modality (which is defined within the \textsf{Iris} 
logic itself), as well as for a wide range of other modalities in the 
literature~\cite{restall2002introduction}. These modalities are typically multi-modal, 
with a form of roughly $M~\overline{ctxt}~\phi\triangleq\ldots\rightarrow \phi~\overline{ctxt}$.

@@

This definition also reveals a subtlety about the denotation of assertions $\phi$ --- they are effectively \emph{predicates} of some kind, functions from some possible world into some logical algebra such as a BI-algebra~\cite{ohearn1999bunched}. 
Taking this interpretation literally gives rise to a convenient way to embed new modalities in \textsf{Iris} 
in a way well-suited to many use cases. 
If $B$ is a type with a BI-algebra structure, then any space of predicate functions (e.g., $\mathsf{val}\rightarrow B$ as a simplification) also carries a BI-algebra structure, lifted from $B$, often called the \emph{pointwise lifting} of the algebra. A number of 
projects for systems software verification~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} now take advantage of this,
explicitly or implicitly lifting the BI-algebra structure from \textsf{Iris} itself to \textsf{Iris} predicates 
used as assertions within modalities. 
Connecting this to \textsf{Iris}'s typeclasses for interactive proofs~\cite{krebbers2018mosel,krebbers-17} then permits working within the embedded logic as in \textsf{Iris}.
